---
title: "Migration Dynamics and Quality of Life Index"

format:
  html:
    code-fold: true
    code-summary: "Code"
    toc: true             
    toc-location: left  
    theme: cosmo        
    embed-resources: true
  pdf:
    colorlinks: true
    papersize: a4
    fontsize: 11pt
    margin-left: 2cm
    margin-right: 2cm
    margin-top: 2cm
    margin-bottom: 2cm
    
execute:
  warning: false
  message: false
---

Welcome to our project page.

**Our Team:**
[Selin DÜZEN](https://emu-hacettepe-analytics.github.io/emu660-spring2026-Selindznn/project.html)

[Burcu BARIŞ](https://emu-hacettepe-analytics.github.io/emu660-spring2026-burcubaris/project.html)

Keep an eye on this space to stay updated with my project activities.

# 1. Project Overview and Scope

Migration is not merely the geographical displacement of individuals; it is also a reflection of economic pursuits, social opportunities, and structural quality in a region. In Turkey, the difference in development and living standards between provinces constitutes one of the primary dynamics of internal migration movements.

In this project, using TÜİK (Turkish Statistical Institute) data, we aim to examine the multidimensional relationship between Net Migration (Net Migration = Migrant Receiving - Migrant Giving) figures for all 81 provinces of Türkiye and the Livability Index indicators in these provinces using analytical methods. Our goal is to reveal, through data analytics and statistical models, which vital factors (economy, education, health, environmental quality, security, etc.) cause people to leave their current cities or move to specific cities.

*Creating Migration Profiles:* To identify the "Net Migration Champions" (provinces receiving the most net migrants) and "Migration Sources" (provinces giving the most net migrants) in Türkiye and to map them geographically.
*The Impact of Livability Dimensions:* To statistically demonstrate which of the sub-components of the livability index (Income and Wealth, Health, Education, Environment, Security, etc.) are more dominant "push" or "pull" forces influencing migration decisions.

# 2. Data


## 2.1 Data Source

The datasets used in the project are: 

1.[Population size of the provinces](https://veriportali.tuik.gov.tr/api/tr/data/downloads?t=i&p=btt%2FI0Npd%2BLvtOvCSZRIUZxbvXE3lIAf5bE6X3CdLQJxPknvezk%2BE6i3Md3GTPYHYMhzGdG2vbYkzOXomic6rY8F%2FyL8NGud1ChYCxoEgy5FkIT%2BMezMLoMVyKJUHwSgAQ%2BDeuZaUgEdXmyanOwSEQ%3D%3D)

2. [Migration received by provinces according to the reason for migration]
(https://veriportali.tuik.gov.tr/api/tr/data/downloads?t=i&p=%2FMyiBh3DmS%2Bv%2BBjCf9%2BzPq%2B96lw8HZTF50uTl9jqBks%2BK5WE6sQNt%2Fqka0J5%2BImWGTGxKxiPcrz7kNG6I3rKkrLi%2B8mZkMkz9reke8%2BHd%2Fo%3D)

3. [Migration rates by provinces according to the reason for migration] (https://veriportali.tuik.gov.tr/api/tr/data/downloads?t=i&p=%2FMyiBh3DmS%2Bv%2BBjCf9%2BzPq%2B96lw8HZTF50uTl9jqBks%2BK5WE6sQNt%2Fqka0J5%2BImWGTGxKxiPcrz7kNG6I3rKkrLi%2B8mZkMkz9reke8%2BHd%2Fo%3D)

4. [Net migration rate for provinces according to reasons for migration] Calculated as "Migration Outgoing - Migration Incoming" 

5. [Life Index Rankings and Index Values by Province] (https://veriportali.tuik.gov.tr/api/tr/data/downloads?t=i&p=5IS1p1Z%2Btvlc0qxFhZzbC9g6NOWXugwdILs8mAMegx5w5sIpUqxydG6%2B5%2FWn1HGIIXJVqd%2Fwn%2B78duD2fsK84VEREA5Sn4QY5oeOWRHZRvVYOaDudiOCwll9hFpmp5W8)

## 2.2 General Information About Data

Our dataset tracks migration metrics across all 81 provinces of Türkiye from 2018 to 2024. It combines raw inbound/outbound migration figures with detailed columns explaining 12 different reasons for migration, cross-references them with the Turkish Statistical Institute's (TÜİK) comprehensive Life Index scores, and shows the relationships between them.

## 2.3 Reason of Choice

We chose this data because migration cannot be explained solely by economic factors. By combining net migration metrics with the multidimensional Life Index (which assesses health, education, security, and environment), we can reveal the relationship between internal migration in Türkiye and life indices.

## 2.4 Preprocessing

The net migration values are calculated using migration data received and given by provinces in different Excel spreadsheets and the char formatted values containing spaces in the data were updated to numeric values.

```{r}
#install.packages("ggplot2") 
#install.packages("dplyr")   
#install.packages("openxlsx")
#install.packages("ggrepel")
library(ggplot2)
library(dplyr)
library(readxl) 
library(writexl) 
library(stringr)
library(forcats)
library(tidyr)
library(openxlsx)
library(ggplot2)
library(readxl)
library(dplyr)
library(ggrepel)
```

### 2.4.1. Calculating net migration values and creating a separate Excel file and checking and correcting numerical values.
```{r}
VerilenGoc <- read_excel("C:/EMU660_Project/emu660-spring2026-Selindznn/VerilenGoc.xlsx", col_names = TRUE)
AlinanGoc  <- read_excel("C:/EMU660_Project/emu660-spring2026-Selindznn/AlinanGoc.xlsx", col_names = TRUE)
toplam_sutun <- ncol(VerilenGoc)
fark_df <- VerilenGoc 
for (i in 3:toplam_sutun) {
  v_clean <- as.character(VerilenGoc[[i]])
  v_clean <- str_replace_all(v_clean, "\\s", "") 
  VerilenGoc[[i]] <- as.numeric(v_clean)        
  a_clean <- as.character(AlinanGoc[[i]])
  a_clean <- str_replace_all(a_clean, "\\s", "")    
  AlinanGoc[[i]] <- as.numeric(a_clean)
}
fark_df[, 3:toplam_sutun] <- VerilenGoc[, 3:toplam_sutun] - AlinanGoc[, 3:toplam_sutun]
for (i in 3:toplam_sutun) {
  fark_df[[i]] <- as.numeric(fark_df[[i]])
}
write_xlsx(fark_df, "C:/EMU660_Project/emu660-spring2026-Selindznn/NetGoc.xlsx")
print(VerilenGoc[, 3:toplam_sutun])
```

### 2.4.2. Calculating net migration rate using population data of provinces and creating an Excel file.
```{r}
kaynak_yolu <- "C:/EMU660_Project/emu660-spring2026-Selindznn/NetGoc.xlsx"
hedef_yolu  <- "C:/EMU660_Project/emu660-spring2026-Selindznn/IlNufuslarii.xlsx"

df_kaynak <- read_excel(kaynak_yolu, col_names = TRUE)

NetGocVerisi <- df_kaynak[1:82, 3, drop = TRUE] 
wb <- loadWorkbook(hedef_yolu)
writeData(wb, sheet = 1, x = "Net Göç", startCol = 3, startRow = 1, colNames = FALSE)
writeData(wb, sheet = 1, x = NetGocVerisi, startCol = 3, startRow = 2, colNames = FALSE)

saveWorkbook(wb, hedef_yolu, overwrite = TRUE)
```
# 3. Analysis

## 3.1 Exploratory Data Analysis

### 3.1.1. Changes in Total Migration Values by Year (2018-2024)
The changes in the total amount of migration over the years were shown graphically.
```{r}
grafik_verisi <- VerilenGoc[VerilenGoc[, 2] == "Toplam-Total", ]
print(grafik_verisi)
head(grafik_verisi)

yil <- factor(grafik_verisi[[1]], levels = sort(unique(grafik_verisi[[1]]), decreasing = TRUE))
goc_miktari  <- as.numeric(grafik_verisi[[3]])/1000  

ggplot(data = grafik_verisi, aes(x = yil, y = goc_miktari)) +
  geom_col(fill = "steelblue", alpha = 0.7) + 
  geom_line(color = "black", size = 2) + 
  geom_point(color = "darkred", size = 3) +
  geom_text(aes(label = round(goc_miktari, 1)), vjust = -0.5, size = 3.5) +
  labs(
    title = "Yıllara Göre Toplam Verilen Göç Miktarı",
    x = "Yıl",
    y = "Göç Miktarı (x1.000)"
  )
```
  
### 3.1.2. Changes in Total Migration Values by Year for Ankara Provience
The graphic shows the changes in total migration amounts over the years for Ankara province.

```{r}
grafik_verisi <- VerilenGoc[VerilenGoc[, 2] == "Ankara", ]
print(grafik_verisi)
head(grafik_verisi)

yil <- factor(grafik_verisi[[1]], levels = sort(unique(grafik_verisi[[1]]), decreasing = TRUE))
goc_miktari  <- as.numeric(grafik_verisi[[3]])/1000  

ggplot(data = grafik_verisi, aes(x = yil, y = goc_miktari)) +
  geom_col(fill = "steelblue", alpha = 0.7) + 
  geom_line(color = "black", size = 2) + 
  geom_point(color = "darkred", size = 3) +
  geom_text(aes(label = round(goc_miktari, 1)), vjust = -0.5, size = 3.5) +
  labs(
    title = "Ankara için Yıllara Göre Toplam Verilen Göç Miktarı",
    x = "Yıl",
    y = "Göç Miktarı (x1.000)"
  )
```

### 3.1.3. Migration Figures by Province in 2024
```{r}
AlinanGoc  <- read_excel("C:/EMU660_Project/emu660-spring2026-Selindznn/AlinanGoc.xlsx",col_names = TRUE)

colnames(AlinanGoc)[1:3] <- c("Yil", "Il", "Goc_Miktari")
veri_2024 <- subset(AlinanGoc, Yil == 2024 & Il != "Toplam-Total")
veri_2024$Goc_Miktari_Bin <- veri_2024$Goc_Miktari / 1000


ggplot(veri_2024, aes(x = Goc_Miktari_Bin, y = reorder(Il, Goc_Miktari_Bin))) +
  geom_col(aes(fill = Goc_Miktari_Bin > 0)) +
  scale_fill_manual(values = c("coral", "skyblue"), guide = "none") +
  theme_minimal() +
  labs(
    title = "2024 Yılı İllerin Alınan Göç Miktarları",
    x = "Alınan Göç Miktarı (Bin Kişi)",
    y = "İller"
  )
```


### 3.1.4. Net Migration Figures of Provinces in 2024

A graphical representation of the calculation made using the previously calculated data given migration-received migration data.

```{r}
NetGoc  <- read_excel("C:/EMU660_Project/emu660-spring2026-Selindznn/NetGoc.xlsx",col_names = TRUE)

colnames(NetGoc)[1:3] <- c("Yil", "Il", "Goc_Miktari")
veri_2024 <- subset(NetGoc, Yil == 2024 & Il != "Toplam-Total")
veri_2024$Goc_Miktari_Bin <- veri_2024$Goc_Miktari / 1000


ggplot(veri_2024, aes(x = Goc_Miktari_Bin, y = reorder(Il, Goc_Miktari_Bin))) +
  geom_col(aes(fill = Goc_Miktari_Bin > 0)) +
  scale_fill_manual(values = c("coral", "skyblue"), guide = "none") +
  theme_minimal() +
  labs(
    title = "2024 Yılı İllerin NEt Göç Miktarları",
    x = "Net Göç Miktarı (Bin Kişi)",
    y = "İller"
  )  
```
### 3.1.5. Graph showing the distribution of migration reasons that make up net migration data.

Migration reasons include: change of job, starting/finding a job, change of marital status/family-related reasons, better housing and living conditions, migration related to any member of the household/family, returning to family home/hometown, health/care, buying a house, retirement, other, and unknown.

```{r}
NetGoc  <- read_excel("C:/EMU660_Project/emu660-spring2026-Selindznn/NetGoc.xlsx",col_names = TRUE)

goc_nedenleri <- c("Tayin/Is Degisikligi", "Ise Baslamak/Is Bulmak", "Egitim", 
                   "Medeni Durum/Ailevi", "Daha Iyi Konut/Yasam", "Haneye Bagimli Goc", 
                   "Memlekete Donus", "Saglik/Bakim", "Ev Alinmasi", 
                   "Emeklilik", "Diger", "Bilinmeyen")

colnames(NetGoc)[1:3] <- c("Yil", "Il", "Toplam_Goc")
colnames(NetGoc)[4:15] <- goc_nedenleri
veri_2024 <- subset(NetGoc, Yil == 2024 & Il != "Toplam-Total")

veri_uzun <- pivot_longer(veri_2024, 
                          cols = 4:15, 
                          names_to = "Goc_Nedeni", 
                          values_to = "Miktar")

veri_uzun$Miktar_Bin <- veri_uzun$Miktar / 1000

goc_grafigi <- ggplot(veri_uzun, aes(x = Miktar_Bin, y = reorder(Il, Toplam_Goc), fill = Goc_Nedeni)) +
  geom_col(position = "stack") +
  theme_minimal() +
  labs(
    title = "2024 Yılı İllerin Göç Nedenlerine Göre Dağılımı",
    subtitle = "Miktarlar bin kişi cinsinden ifade edilmiştir",
    x = "Net Göç Miktarı (Bin Kişi)",
    y = "İller",
    fill = "Göç Nedenleri"
  ) +
  theme(
    legend.position = "bottom",
    legend.text = element_text(size = 8),
    axis.text.y = element_text(size = 7)
  ) +
  guides(fill = guide_legend(ncol = 3))
  
print(goc_grafigi)
```

## 3.2 Trend Analysis

The graph shows the relational analysis between migration rate data and life indices of the provinces.

```{r}
yasam_df <- read_excel("C:/EMU660_Project/emu660-spring2026-Selindznn/YasamEndeksi.xlsx", col_names = TRUE)
goc_df <- read_excel("C:/EMU660_Project/emu660-spring2026-Selindznn/Gochizi.xlsx", col_names = TRUE)

colnames(yasam_df)[1:2] <- c("Il", "Endeks")
colnames(goc_df)[1:3]   <- c("Yil", "Il", "Goc")

goc_2024 <- subset(goc_df, Yil == 2024 & Il != "Toplam-Total")

veri_seti <- inner_join(yasam_df, goc_2024, by = "Il")

grafik <- ggplot(veri_seti, aes(x = Endeks, y = Goc, label = Il)) + 
  geom_point(color = "#1f77b4", size = 3, alpha = 0.7) + 
  geom_text_repel(size = 3, max.overlaps = 15) + 
  geom_smooth(method = "lm", color = "red", se = TRUE, formula = y ~ x) + 
  labs(
    title = "İl Yaşam Endeksi ve 2024 Net Göç Hızı İlişkisi",
    x = "Genel Yaşam Endeksi",
    y = "Net Göç Hızı (‰)"
  ) +
  theme_minimal()

print(grafik)

```

## 3.4 Results

Our time series analysis (2018–2024) shows that overall migration numbers fluctuate significantly based on national economic shifts and major external events (e.g., pandemics, earthquakes). When we looking at major metropolitan areas like Ankara, we see a distinct pattern of high incoming and high outgoing figures, reflecting its status as a dynamic political and educational hub. 
Our 2024 province-by-province breakdown confirms that the geographic distribution of migration in Turkey is highly polarized. A small cluster of economically dominant provinces acts as "Net Migration Champions," while a vast majority of eastern and rural provinces serve as continuous "Migration Sources. 
The stacked column chart for 2024 reasons for migration provides crucial insights. Rather than simple employment changes ("Tayin/İş Değişikliği"), a massive portion of net migration is driven by:Education ("Eğitim"): Causing heavy youth mobility toward university cities.Better Housing and Living Conditions ("Daha İyi Konut/Yaşam") Indicating that families are moving reactively to secure better urban standards.Dependent Migration ("Haneye Bağımlı Göç")Showing that migration is rarely an individual choice but a household-level transition. 
Our trend analysis using geom_smooth(method = "lm") reveals a positive correlation between a province's General Life Index and its Net Migration Rate.


# 4. Results and Key Takeaways

As the Life Index of a province increases, its net migration rate trends upward. This statistically demonstrates that safety, better healthcare, environmental quality.  

The positive trend line between the Life Index and Net Migration proves that modern citizens actively seek out "livability"—safer streets, greener spaces, and better educational institutions for their children. Cities with high Life Indexes are facing massive population influxes. If left unmanaged, this will cause severe strain on metropolitan infrastructures (traffic, housing crises, environmental degradation). Meanwhile, some provinces lose their young, educated workforce (internal brain drain), making it harder for them to improve their regional development scores.

Governments should focus on raising the localized "Life Index" components—such as building better hospitals, upgrading schools, and enhancing civic safety in negative-net-migration provinces—to encourage citizens to stay.

